A Constraint Solving Approach to Parikh Images of Regular Languages

### 2022-07-08 main.tex
Fitting till preliminaries

Since Parikh images are semilinear, any Parikh image can be written as a set of linear
equations. The following construction for an NFA $\Automaton = \AutomatonTuple$ is based on the version for context-free grammars presented in \cite{generate-parikh-image}:
\begin{equation}
\begin{aligned}
\ParikhMap(\Automaton) := 
& \AndComp{\Letter \in \Alphabet}{\LetterVar_{\Letter} = \sum_{\Transition = \FromLabelTo{}{\Letter}{} \in \Transitions} \TransitionVar_{\Transition}
}\\
& \AndComp{\State \in \AcceptingStates}{\FinalStateVar_{\State} > 0 \longrightarrow  \StateVar_{\State} > 0}\\
& \AndComp{\State \in \States}{
  \left(\FinalStateVar_{\State} \text{ if } \State \in \AcceptingStates \text{ otherwise } 0 \right) +
  \sum_{\Transition = \FromLabelTo{\State}{}{}} \Filter(\Transition) - \sum_{\Transition = \FromLabelTo{}{}{\State}} \Filter(\Transition)
= \begin{cases}
    1 \text{  if $\State = \InitialState$} \\
    0 \text{ otherwise}
  \end{cases}
}\\
& \AndComp{\Transition = \FromLabelTo{\State'}{}{\State} \in \Transitions}{
  \TransitionVar_{\Transition} > 0 \longrightarrow \StateVar_{\State} > 0
} \\
& \AndComp{\State \in F}{
  \StateVar_{\State} > 0 \longrightarrow \FinalStateVar_{\State} = 1
} \\
&\AndComp{\State \in \States}{
  \StateVar_{\State} > 0 \longrightarrow
  \OrComp{\Transition = \FromLabelTo{\State}{}{\State'} \in \Transitions}{
    \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
    \TransitionVar_{\Transition} \geq 1 \land
  \StateVar_{\State'} \geq 1
    }
}
\end{aligned}
\label{eq:generate-parikh}
\end{equation}

@@

\section{Projections on Parikh Maps}\label{sec:generalised}

It is easy to see that the Parikh map~$\ParikhMap$ represents a
homomorphism from the (free) non-commutative monoid~$\Strings$ to the
(free) commutative monoid~$\Naturals^k$. As we are often not
interested in the full Parikh image, but only in some projections of
it, for the purposes of this paper we use a generalised version of the
Parikh map. We consider \emph{arbitrary} homomorphisms
$\Map:\: \Sigma^* \to \Monoid$, where
$\Monoid = \left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative
monoid; this means that $\Map(\epsilon) = 0_{\Monoid}$ and
$\Map(u \Concat w) = \Map(u) \MonoidOp \Map(w)$. We give several
examples of such generalised Parikh image later in this section.

>>>>

Since Parikh images are semilinear, any Parikh image can be written as a set of linear
equations. The following construction for an NFA $\Automaton = \AutomatonTuple$ is based on the version for context-free grammars presented in \cite{generate-parikh-image}:
\begin{equation}
\begin{aligned}
\ParikhMap(\Automaton) := 
& \AndComp{\Letter \in \Alphabet}{\LetterVar_{\Letter} = \sum_{\Transition = \FromLabelTo{}{\Letter}{} \in \Transitions} \TransitionVar_{\Transition}
}\\
& \AndComp{\State \in \AcceptingStates}{\FinalStateVar_{\State} > 0 \longrightarrow  \StateVar_{\State} > 0}\\
& \AndComp{\State \in \States}{
  \left(\FinalStateVar_{\State} \text{ if } \State \in \AcceptingStates \text{ otherwise } 0 \right) +
  \sum_{\Transition \in \FromLabelTo{\State}{}{}} \Filter(\Transition) - \sum_{\Transition \in \FromLabelTo{}{}{\State}} \Filter(\Transition)
= \begin{cases}
    1 \text{  if $\State = \InitialState$} \\
    0 \text{ otherwise}
  \end{cases}
}\\
& \AndComp{\Transition = \FromLabelTo{\State'}{}{\State} \in \Transitions}{
  \TransitionVar_{\Transition} > 0 \longrightarrow \StateVar_{\State} > 0
} \\
& \AndComp{\State \in F}{
  \StateVar_{\State} > 0 \longrightarrow \FinalStateVar_{\State} = 1
} \\
&\AndComp{\State \in \States}{
  \StateVar_{\State} > 0 \longrightarrow
  \OrComp{\Transition = \FromLabelTo{\State}{\Label}{\State'} \in \Transitions}{
    \StateVar_{\State} = \StateVar_{\State'} + 1 \land 
    \TransitionVar_{\Transition} \geq 1 \land
  \StateVar_{\State'} \geq 1
    }
}
\end{aligned}
\label{eq:generate-parikh}
\end{equation}

@@

\subsection{Theorem Proving}

We assume basic familiarity with theorem provers, in particular sequent calculus
or similar. For reference, see e.g. \cite{Fitting96a}.

\section{Projections on Parikh Images}\label{sec:generalised}

It is easy to see that the Parikh map~$\ParikhMap$ represents a
homomorphism from the (free) non-commutative monoid~$\Strings$ to the
(free) commutative monoid~$\Naturals^k$. As we are often not
interested in the full Parikh image, but only in some projections of
it. We consider \emph{arbitrary} homomorphisms
$\Map:\: \Sigma^* \to \Monoid$, where
$\Monoid = \left(X;\MonoidOp;0_{\Monoid}\right)$ is a commutative
monoid. We give several
examples of such projections on Parikh images later in this section.
